Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2021-08-24 article/draft.tex
add Beatrice's point about the natural disappearance of data

Although it is currently impossible to verify permanent deletions using social media data, their effects can be observed using an adapted methodology.
Following the Capitol riots in January 2021, Twitter has announced the suspension of more than $70$ $000$ accounts related to QAnon.
\footnote{\url{https://blog.twitter.com/en_us/topics/company/2021/protecting--the-conversation-following-the-riots-in-washington--}} 
A journalist has recently investigated the effects of this measure by counting the number of times some hashtags specific to the QAnon movement have been used over time, such as \#QAnon or \#WWG1WGA.
A drastic reduction in hashtag use was found when comparing the first quarter of $2021$ with the first quarter of $2021$.
\footnote{\url{https://twitter.com/Shayan86/status/1394742784683298818}}
Thus, counting specific hashtag use over time allows to indirectly observe permanent suspensions on Twitter, and there is a need for similar methodologies to be experimented on the other platforms' data.

@@

How to best tackle online misinformation is an active area of research, and developing effective interventions against misinformation depends on understanding the psychology of interacting with social media (Pennycook et Rand 2021~\cite{pennycook2021psychology}). 
Adding fact-checking labels to controversial content is certainly the intervention whose effectiveness was the most studied so far. 
It was shown that credibility indicators do reduce misperceptions and sharing, although their impact varied vastly depending on the presentation of the warnings (Yaqub 2020~\cite{yaqub2020effects}, Sharevski 2021~\cite{sharevski2021misinformation}).
Moreover, a fact-checking label can lead to unpredicted consequences depending on the perception of the person reading it.
On Facebook, Twitter and YouTube, fact-checking messages are currently applied only to some content, while the rest of the platform remains unlabelled. 
Such sparce application of warnings can lead to an `implied truth' effect where users may assume that content without warnings have actually been verified and shown to be true (Pennycook et al. 2020~\cite{pennycook2020implied}).
Using another approach, Mosleh et al. (\cite{mosleh2021perverse}) identified Twitter users sharing false news and replied to their false tweets with links to fact-checking websites.
They observed a decrease in quality and accuracy of the content subsequently shared by the corrected user. 
These works highlight the necessity to investigate potential unintended effects of well-meaning interventions.
The methods presented in this paper can help to monitor not only the platforms' actions against misinformations, but also their subsequent consequences in real-life settings.

>>>>

Although it is currently impossible to verify permanent deletions using social media data, their effects can be observed using an adapted methodology.
Following the Capitol riots in January 2021, Twitter has announced the suspension of more than $70$ $000$ accounts related to QAnon.
\footnote{\url{https://blog.twitter.com/en_us/topics/company/2021/protecting--the-conversation-following-the-riots-in-washington--}} 
A journalist has recently investigated the effects of this measure by counting the number of times some hashtags specific to the QAnon movement have been used over time, such as \#QAnon or \#WWG1WGA.
A drastic reduction in hashtag use was found when comparing the first quarter of $2021$ with the first quarter of $2021$.
\footnote{\url{https://twitter.com/Shayan86/status/1394742784683298818}}
Thus, counting specific hashtag use over time allows to indirectly observe permanent suspensions on Twitter, and there is a need for similar methodologies to be experimented on data from other platforms.
However, a point of concern is that data can also disappear because of other factors than the interventions of platforms.
As an example, in 2012, McMinn et al. \cite{mcminn2013building} collected 120 millions of Tweets on very general topics. 
The published set of Tweet IDs was then used to re-run the same collection in 2018, but only 67 millions of Tweets (55\%) were retrieved, probably due to tweets beeing removed by users and Twitter accounts being closed (Mazoyer et al. 2020~\cite{mazoyer2020french}).
As data naturally disappear from platforms over time, it hard to identify whether platforms or users are the ones responsible for their removal. 
Even in the case of the QAnon accounts, it is hard to say how much of the drop in hashtag use is due to the Twitter intervention or to accounts moving to other platforms out of fear of being shut down.

@@

How to best tackle online misinformation is an active area of research, and developing effective interventions against misinformation depends on understanding the psychology of interacting with social media (Pennycook et Rand 2021~\cite{pennycook2021psychology}). 
Adding fact-checking labels to controversial content is certainly the intervention whose effectiveness was the most studied so far. 
It was shown that credibility indicators do reduce misperceptions and sharing, although their impact varied vastly depending on the presentation of the warnings (Yaqub 2020~\cite{yaqub2020effects}, Sharevski 2021~\cite{sharevski2021misinformation}).
Moreover, a fact-checking label can lead to unpredicted consequences depending on the perception of the person reading it.
On Facebook, Twitter and YouTube, fact-checking messages are currently applied only to some content, while the rest of the platform remains unlabelled. 
Such sparce application of warnings can lead to an `implied truth' effect where users may assume that content without warnings have actually been verified and shown to be true (Pennycook et al. 2020~\cite{pennycook2020implied}).
Using another approach, Mosleh et al. \cite{mosleh2021perverse} identified Twitter users sharing false news and replied to their false tweets with links to fact-checking websites.
They observed a decrease in quality and accuracy of the content subsequently shared by the corrected user. 
These works highlight the necessity to investigate potential unintended effects of well-meaning interventions.
The methods presented in this paper can help to monitor not only the platforms' actions against misinformations, but also their subsequent consequences in real-life settings.
